MIMO Channel Shaping and Rate Maximization\\Using Beyond-Diagonal RIS

### 2025-12-09 review/response_250823.tex

\documentclass[
	fontsize=11pt,
	paper=a4,
	foldmarks=false
]{scrartcl}

\input{../manuscript/preamble}
\usepackage{response}

\setkomavar{signature}{Yang Zhao, Hongyu Li, Bruno Clerckx, and Massimo Franceschetti}
\setkomavar{date}{\today}
\setkomavar{subject}{Response to Decision on Manuscript T-SP-33360-2025.R1}

\begin{document}
\begin{letter}{%
		Prof. Wei Yi\\
		Associate Editor\\
		IEEE Transactions on Signal Processing
	}
	\opening{Dear Editor and Reviewers,}
	Thank you for giving us an opportunity to revise the paper ``MIMO Channel Shaping and Rate Maximization Using Beyond-Diagonal \gls{ris}''.
	Your feedback and suggestions have been invaluable in helping us improve the quality of the manuscript.
	Below we prepare a point-to-point response and highlight the corresponding changes in text, where labels have been matched to the submission for your convenience.
	We hope the revisions and clarifications make the manuscript meet the TSP publication standards.
	\closing{Yours sincerely,}
\end{letter}

\begin{editor}
	\summary{
		In this round of review, while two reviewers have acknowledged the improvements made in the revised paper, the third reviewer has raised critical concerns regarding the practical usefulness and overall contribution of the work. Specifically, the reviewer points out that the bounds on the singular values of \gls{bd}-\gls{ris} may not be easily exploitable in practice, and that the number of derived inequalities appears excessive without demonstrating clear practical value. Furthermore, the reviewer notes that the proposed method is limited to single-user scenarios without QoS constraints, thereby limiting its novelty and relevance when compared to existing literature.
	}

	\reply{
		Please refer to our response to Reviewer 3.
	}
\end{editor}

\begin{reviewer}
	\summary{
		(There are no comments.)
	}
\end{reviewer}

\begin{reviewer}
	\summary{
		This version has addressed all my comments.
	}

	\reply{
		Thank you for your positive feedback and continued support.
	}
\end{reviewer}

\begin{reviewer}
	\summary{
		The reviewer sincerely appreciates the authors’ efforts in revising the manuscript and preparing the response. However, from the reviewer’s viewpoint, the major issue remains unresolved --- the major results in the manuscript seem not quite useful.
	}

	\comment{
	The reviewer does not think the bounds on singular values of \gls{bd}-\gls{ris} (Corollary \ref{co:nd_sv_prod_subset} - \ref{co:nd_capacity_snr_general}) are useful. In fact, as reflected in \eqref{iq:sv_3_individual} - \eqref{iq:sv_3_ordering}, these bounds are actually not easily exploitable. Especially, recalling \gls{bd}-\gls{ris} practically has a number of elements no smaller than several tens, the inequalities seem to be too many to be useful.
	}

	\reply{
		It is a pity that the reviewer found our results not quite useful. We believe there could a misunderstanding on the main gist and logic flow of the paper. And we will try to explain this as follows.
		\begin{enumerate}
			\item We are not dealing with ``bounds on singular values of \gls{bd}-\gls{ris}''. The research question is using \gls{bd}-\gls{ris} to shape \gls{mimo} in terms of \emph{channel} singular values and their functions. We believe a comprehensive answer serves as a theoretical support/reference for the vast number of \gls{ris} research papers and real-world applications.
			\item The reviewer may have reasoned that ``\eqref{iq:sv_3_individual} - \eqref{iq:sv_3_ordering} are not easily exploitable'' \textrightarrow ``Corollary \ref{co:nd_sv_prod_subset} - \ref{co:nd_capacity_snr_general} are not useful'' \textrightarrow ``the major results in the manuscript seem not quite useful.'' However, the message we were trying to convey is that Corollary \ref{co:nd_sv_prod_subset} can be hard to exploit so that Corollaries \ref{co:nd_sv_prod_tail} -- \ref{co:nd_capacity_snr_extreme} are introduced to show its profound implications and \eqref{iq:sv_3_individual} - \eqref{iq:sv_3_ordering} are introduced to visualize its complications. Then we develop optimization methods to validate those bounds and handle more general singular value functions. We believe the most important results are the propositions, the closed-form \gls{bd}-\gls{ris} solutions for different shaping objectives, and the geodesic \gls{bd}-\gls{ris} design algorithm.
			\item We believe these theoretical results are useful in the sense that they provide strong mathematical supports to \emph{understand exactly} to what extent passive \gls{ris} can shape the \gls{mimo} channel. The research question is answered from different perspectives in Propositions \ref{pp:dof} -- \ref{pp:nd}. It turns out that Proposition \ref{pp:nd} induces Corollary \ref{co:nd_sv_prod_subset} with exceeding number of inequalities as indicated by the reviewer; but they indeed apply to arbitrary number of \gls{ris} elements. We then carefully pick \emph{subsets} of those in Corollaries \ref{co:nd_sv_prod_tail} -- \ref{co:nd_capacity_snr_extreme} to provide (i) ready-to-use performance bounds for typical wireless applications and (ii) closed-form optimal \gls{ris} solutions. In the context of \gls{bd}-\gls{ris}-aided \gls{mimo}, those corollaries translate to:
			\begin{itemize}
				\item closed-form transmit precoder for spatial multiplexing with a limited number of \gls{rf} chains;
				\item closed-form transmit precoder for \gls{mimo} wireless power transfer with \gls{rf} combining;
				\item maximum and minimum channel power gains, which further motivates low-complexity transmit precoder design for wireless communication and power transfer in Section \ref{sc:rate};
				\item channel capacity at general and extreme \glspl{snr} with water-filling precoder.
			\end{itemize}
			The above non-exhaustive list demonstrates the outcomes of understanding channel shaping limits; we really hope our ingenious readers can discover more results specific to their research.
			In fact, that is the reason we presented Proposition \ref{pp:nd} and Corollary \ref{co:nd_sv_prod_subset} as general as possible.
		\end{enumerate}

		The manuscript has been updated as follows.

		\change{
			Unlike existing works that focus on specific performance metrics or deployment scenarios, we aim for an understanding of the theoretical shaping limits (via analysis) and the achievable shaping results (via optimization) that are broadly applicable across diverse wireless applications.
			We believe a comprehensive shaping answer can serve as a theoretical support/reference for the vast number of \gls{ris} research papers and real-world applications.
			Without a framework that identifies the fundamental shaping limits of \gls{ris}, the design of truly optimal and efficient architectures will remain elusive.
		}

		\change{
			Corollary~\ref{co:nd_sv_prod_subset} applies to arbitrary number of \gls{ris} elements as inherited from Proposition~\ref{pp:nd}.
			The set \eqref{iq:horn}, also recognized as a variation of Horn's inequality \cite{Bhatia2001}, provides a comprehensive analytical answer to the shaping question -- it can be interpreted as the \emph{outer bounds} of the achievable singular value region of the \gls{bd}-\gls{ris}-aided \gls{mimo} channel.
			An example is given by \eqref{iq:sv_3_bounds} and their visualization in Fig.~\ref{fg:singular_region}.
			Remarkably, the number of inequalities in \eqref{iq:horn} increases exponentially with $N_\mathrm{S}$.
			At a first glance the results may seem excessive to be useful; but they are given in this form to be general and one can pick any \emph{subset} of them for specific applications.
			% , but in practice we can use only a subset of them that are relevant to typical implementations, as demonstrated in the paper.
			% While finding direct applications for every resulting bound is challenging, they together provide a rich foundation on the limits of channel shaping.
			% provide a rich theoretical foundation and performance limit for channel shaping using \gls{bd}-\gls{ris}.
			% One may choose any subset of those for specific wireless applications.
			% Below we showcase some useful results in Corollaries \ref{co:nd_sv_prod_tail} -- \ref{co:nd_capacity_snr_extreme} to demonstrate its profound implications.
			Below we showcase how to induce some ready-to-use wireless performance bounds with closed-form \gls{bd}-\gls{ris} solutions from Corollary~\ref{co:nd_sv_prod_subset}.
			% pick \emph{subsets} of \eqref{iq:horn} to induce some ready-to-use wireless performance bounds with closed-form \gls{bd}-\gls{ris} solutions.
			The applications mentioned therein are non-exhaustive; we really hope our ingenious readers can discover more results specific to their research.
		}

		\change{
			Proposition~\ref{pp:dof} -- \ref{pp:nd} and the resulting Corollaries provide a partial answer to the channel shaping question.
			% Those analyses are valuable since they provide a theoretical foundation for one to understand exactly \emph{to what extent} and \emph{how} a \gls{bd}-\gls{ris} can shape the \gls{mimo} channel in terms of simple singular value functions.
			These analyses provide a theoretical foundation for understanding exactly \emph{how} and \emph{to what extent} a \gls{bd}-\gls{ris} can shape the \gls{mimo} channel regarding typical singular value functions under specific channel conditions.
			Extending the results to more general setups is challenging due to the limited branch matching capability and the inherent tradeoff in mode alignment.
		}

	}

	\comment{
	The rest main results are Alg. \ref{ag:rcg} and rate optimization in Sec. \ref{sc:rate}. However, these algorithms only apply to simple scenarios, e.g., single user case, without additional constraints (QoS, interference, CRB). In fact, rate maximization for single user have been extensively studied by existing literature. There exist a huge body of works on \gls{bd}-\gls{ris} beamforming design considering much more complicated scenarios, e.g., multi-user context and scenarios with more complicated constraints. See the survey \cite{Li25x}.
	}

	\reply{
		We appreciate and fully respect the reviewer's feedback and suggestions.
		In fact, the survey paper \cite{Li25x} made its debut 10 months after this paper and has heavily cited our algorithms.
		We choose not to do the other way around.
		It is also clear from \cite[Table~V]{Li25x} that other \gls{mimo} rate maximization designs are quite limited even in the single-user scenario -- they considered either fully-connected \gls{bd}-\gls{ris} or line-of-sight direct channel, both of which fall into special cases of our channel shaping analysis in Section \ref{sc:shaping_analytical}.
		That is, it remains unclear in the literature how to design a symmetric/asymmetric \gls{bd}-\gls{ris} of arbitrary group size for rate maximization under general \gls{mimo} channel conditions.
		% Simulation results in Section \ref{sc:simulation} demonstrate that t
		The proposed geodesic manifold optimizer not only fills in this gap but also features higher computational efficiency than the canonical off-the-shelf optimizer, as analyzed in Section \ref{sc:shaping_numerical} and simulated in Section \ref{sc:simulation}.
		Besides, this paper is a pioneer to demonstrate how channel shaping can be applied to rate maximization and other joint beamforming problems, which offers a low-complexity candidate to the de facto alternating optimization.

		On the other hand, we would also like to emphasize that our focus on the single-user case is a deliberate and methodical choice.
		We believe in a ``walk-before-run'' approach.
		A thorough understanding of the fundamental limits of point-to-point \gls{mimo} channel shaping via \gls{ris} is a prerequisite for properly tackling the more complex, resource-limited multi-user scenarios.
		% The primary goal of this paper is to clearly establish the baseline understandings, upon which the
		% It is essential to first establish a solid understanding of the \gls{ris} shaping limits in fundamental point-to-point \gls{mimo}, which is still an open problem in itself, before tackling the significantly more complex multi-user case.
		Our paper aims to provide this foundational baseline, which we see as critical and necessary for future works in the area.
		While Algorithm \ref{ag:rcg} and the two solutions in Section~\ref{sc:rate} are readily extendable to weighted sum-rate maximization and leakage interference minimization in \gls{mimo} interference channel, we have decided not to include the results in this paper to avoid blurring the focus and confusing our readers.
		We have updated the conclusion to address this concern.
		\change{
			This paper investigated the capability of \gls{bd}-\gls{ris} to shape a \gls{mimo} channel in terms of singular values and their functions.
			We started from a toy example and derived some analytical bounds (with closed-form solutions) on the channel singular values, power gain, and capacity.
			An efficient framework was then proposed to optimize the \gls{bd}-\gls{ris} for a broader class of singular value functions.
			% We focus on a \gls{bd} architecture that allows elements within the same group to interact, enabling more sophisticated signal processing than \gls{d}-\gls{ris}.
			% This translates to a wider dynamic range of and better tradeoff between singular values, resulting in significant power and rate gains.
			% Analytical singular value bounds are derived under typical \gls{ris} deployment scenarios and the Pareto frontiers are characterized via an efficient \gls{rcg} method.
			We also presented two beamforming designs for the rate maximization problem, one for optimal performance and the other exploits shaping implications for much lower complexity while remaining close-to-optimal.
			Extensive simulation show that the significant power and rate gains of \gls{bd}-\gls{ris} over \gls{d}-\gls{ris} stems from its superior \gls{mimo} branch matching and mode alignment potentials, which scales with the number of elements, group size, and \gls{mimo} dimensions.

			The analysis and optimization methods in this paper have been tailored for group-connected \gls{bd}-\gls{ris}.
			Extension to other \gls{ris} architectures remains a promising area for future research.
			One straightforward extension to the multi-sector model \cite{Li2023c} is to retrieve the optimal scattering matrix for each sector individually by methods in this paper and then play with the power splitting factors.
			Meanwhile, transitioning from single- to multi-layer \gls{ris} models \cite{An23b} mirrors that from single- to multi-hop \gls{af} relays; interested readers may be inspired by \cite{Rong2009a}.

			Finally, we remark that the principle of channel shaping is not limited to point-to-point \gls{mimo}.
			Algorithm \ref{ag:rcg} and the two solutions in Section~\ref{sc:rate} are readily extendable to weighted sum-rate maximization and leakage interference minimization in \gls{mimo} interference channel.
			% Those are not included here due to page limits; please refer to our \href{https://arxiv.org/pdf/2407.15196}{arXiv} and \href{https://github.com/snowztail/channel-shaping}{GitHub} version for details.

		}

		% Unfortunately they won't go with the current submission due to the page limits.

		% \begin{subsection}{MIMO Interference Channel Shaping}
		% 	\label{sc:mimo_ic}
		% 	We also consider a \gls{bd}-\gls{ris}-aided \gls{mimo} interference channel of $K$ transceiver pairs where each transmitter and receiver has $N_\mathrm{T}$ and $N_\mathrm{R}$ antennas, respectively, and the \gls{bd}-\gls{ris} has $N_\mathrm{S}$ scattering elements.
		% 	This configuration is denoted as $(N_\mathrm{T} \times N_\mathrm{S} \times N_\mathrm{R})^K$.
		% 	Let $\mathbf{H}_\mathrm{D}^{(kj)} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{B}^{(k)} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{S}}$, $\mathbf{H}_\mathrm{F}^{(j)} \in \mathbb{C}^{N_\mathrm{S} \times N_\mathrm{T}}$ denote the direct channel from transmitter $j$ to receiver $k$, the backward channel of receiver $k$, and the forward channel of transmitter $j$, respectively, where $(j,k) \in [K]^2$.
		% 	Assume all transmitter-\gls{ris}-receiver paths share the same \gls{bd}-\gls{ris} scattering matrix $\mathbf{\Theta}$.
		% 	The equivalent channel from transmitter $j$ to receiver $k$ is
		% 	\begin{equation}
		% 		\label{eq:channel_interference}
		% 		\mathbf{H}^{(kj)} = \mathbf{H}_\mathrm{D}^{(kj)} {+} \mathbf{H}_\mathrm{B}^{(k)} \mathbf{\Theta} \mathbf{H}_\mathrm{F}^{(j)} = \mathbf{H}_\mathrm{D}^{(kj)} {+} \sum_g \mathbf{H}_{\mathrm{B},g}^{(k)} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g}^{(j)},
		% 	\end{equation}
		% 	where $\mathbf{H}_{\mathrm{B},g}^{(k)} \in \mathbb{C}^{N_\mathrm{R} \times L}$ and $\mathbf{H}_{\mathrm{F},g}^{(j)} \in \mathbb{C}^{L \times N_\mathrm{T}}$ are associated with \gls{ris} group $g$, corresponding to the $(g{-}1)L{+}1$ to $gL$ columns of $\mathbf{H}_{\mathrm{B}}^{(k)}$ and rows of $\mathbf{H}_{\mathrm{F}}^{(j)}$, respectively.
		% 	On top of \eqref{eq:channel_interference}, the achievable rate of transmission $k$ is
		% 	\begin{equation}
		% 		R_k = \log \det \biggl(\mathbf{I} + \mathbf{W}_k {\mathbf{H}^{(kj)\mathsf{H}}} \mathbf{Q}_k^{-1} {\mathbf{H}^{(kk)}} \mathbf{W}_k\biggr),
		% 	\end{equation}
		% 	where $\mathbf{W}_k$ is the precoder at transmitter $k$ and $\mathbf{Q}_k = \sum_{j \ne k} {\mathbf{H}^{(kj)}} \mathbf{W}_j \mathbf{W}_j^\mathsf{H} {\mathbf{H}^{(kj)\mathsf{H}}} + \eta \mathbf{I}$ is the interference-plus-noise covariance matrix at receiver $k$.
		% 	The \gls{wsr} maximization problem for \gls{bd}-\gls{ris}-aided \gls{mimo} interference channel is formulated as
		% 	\begin{maxi!}
		% 	{\scriptstyle{\mathbf{\Theta}, \{\mathbf{W}_k\}_{k \in [K]}}}{\sum_{k=1}^K \rho_k R_k}{\label{op:wsr}}{\label{ob:wsr}}
		% 	\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g}{}{}
		% 	\addConstraint{\lVert \mathbf{W}_k \rVert _\mathrm{F}^2 \le P_k. \quad \forall k}{}{}
		% \end{maxi!}
		% 	where $\rho_k \ge 0$ is the weight associated with transmission $k$.
		% 	This non-convex problem can be solved by extending both solutions covered in Section \ref{sc:rate} as detailed below.

### 2025-12-09 manuscript/main.tex
:triangular_flag_on_post: TSP 250823

	Although above works offered inspiring glimpses into the channel shaping potential of passive \gls{ris}, they neither provided in-depth theoretical analysis nor characterized the achievable singular value region.
	Most works \cite{ElMossallamy2021,Meng2023,Huang2023,Chae2023,Wu2019,Santamaria2023a} have also been confined to the conventional \gls{d} architecture where each \gls{ris} element is connected to a dedicated impedance and functions independently of the others, namely, the wave impinging on one element is entirely scattered by itself.
	This architecture is modeled by a diagonal scattering matrix with unit-magnitude diagonal entries that ideally applies a phase shift to the incident signal.
	The idea was soon extended to \gls{bd}-\gls{ris} with group-connected architecture that connects elements within the same group via passive reconfigurable circuit components \cite{Shen2020a} that can be symmetric (e.g., capacitors and inductors) or asymmetric (e.g, ring hybrids and branch-line hybrids \cite{Ahn2006}).
	As such, the wave impinging on one element is able to propagate within the circuit and depart partially from the others of the same group.
	It can thus manipulate both amplitude and phase of the scattered wave while remaining globally passive.
	The main manufacturing complexity of \gls{bd}-\gls{ris} lies in the design and implementation of the circuit network.
	Fortunately, novel topologies such as tree- and forest-connections have been proposed to reduce the number of components for a flexible cost-performance tradeoff \cite{Nerini2024}.
	Other practical challenges such as channel estimation \cite{Li2024}, mutual coupling \cite{Li2023f}, wideband modelling \cite{Li2024a}, multi-sector coverage \cite{Li2023c}, and hardware implementation \cite{Tapie25e} have also been studied in recent literature.
	\gls{bd}-\gls{ris} has been proved to achieve higher spectral efficiency than \gls{d}-\gls{ris} and higher energy efficiency than active \gls{ris} and \gls{af} relay \cite{Santamaria2023,Fang2023,Zhou2023}.
	However, the interplay between \gls{bd}-\gls{ris} and \gls{mimo} is still at infancy stage and the potential remains largely unexplored.
	% Most relevant works are confined to the scenario
	For example, the rate maximization problem \cite{Bartoli2023} has only been tacked in the special case where the direct channel is negligible and the \gls{bd}-\gls{ris} is fully-connected.
	Under those conditions, the mathematical modeling of \gls{bd}-\gls{ris} coincides with that of \gls{af} relay of unit power, although the operation mechanism and noise characteristics are clearly distinct.

	When it comes to signal processing, existing works have mainly invoked the {quasi-Newton} method \cite{Shen2020a}, the \gls{pdd} method \cite{Zhou2023}, and the generic (i.e., non-geodesic) \gls{rcg} method \cite{Li2023c} for the optimization of \gls{bd}-\gls{ris}.
	The first solves an unconstrained problem and projects the solution back to the feasible domain without optimality guarantee.
	The second alternates between the primal variables in the inner layer and the penalty coefficient in the outer layer.
	It is often used to tackle coupled constraints (e.g., \gls{sinr} thresholds under active and passive beamforming) and can be computationally expensive (e.g., $\mathcal{O}(N_{\mathrm{s}}^2)$ for \gls{d}-\gls{ris} and $\mathcal{O}(N_{\mathrm{s}}^4)$ for fully-connected \gls{bd}-\gls{ris}) \cite[Table I]{Zhou2023}.
	The third applies the conjugate gradient method on generic Riemannian manifolds.
	Each iteration consists of an addition on the tangent space and a retraction to the feasible domain, which constitutes a zigzag path departing from and returning to the manifold.
	However, none of them effectively exploits the special structure of \gls{bd}-\gls{ris} for accelerated convergence.

	% \begin{subsection}{Contributions}
	\label{sc:contributions}
	This paper is motivated by a fundamental question:
	\emph{What is the channel shaping capability, in terms of singular values and their functions, of a passive \gls{ris} in \gls{mimo} channels?}
	Unlike existing works that focus on specific performance metrics or deployment scenarios, we aim for an understanding of the theoretical shaping limits (via analysis) and the achievable shaping results (via optimization) that are broadly applicable across diverse wireless applications.
	The contributions of this paper are summarized below.

@@

	Fifth, we validate the analytical bounds and the numerical methods by extensive simulation.
	It is concluded that:
	\begin{itemize}
		\item \gls{bd}-\gls{ris} can widen the dynamic range of channel singular values for enhanced rate, power, and physical-layer security;
		\item The shaping benefits of \gls{bd}-\gls{ris} over \gls{d}-\gls{ris} scale with the number of elements, group size, and \gls{mimo} dimensions;
		% \item The optimal shaping solutions for communication and power transfer coincide when direct channel is negligible and the \gls{bd}-\gls{ris} is fully-connected;
		% \item The optimally shaped \gls{mimo} channel for communication and power transfer\footnote{with \gls{rf} combining} coincide with each other
		% power-gain optimal, rate-optimal, and power-transfer \gls{bd}-\gls{ris} scattering matrix coincide
		% \item A fully-connected \gls{bd}-\gls{ris} may be designed in closed-form for simultaneous optimality of \gls{mimo} communication and power transfer;
		\item \gls{bd}-\gls{ris} can activate multi-stream transmission and achieve the asymptotic \gls{dof} at lower transmit power than \gls{d}-\gls{ris};
		\item The rate gap between the \gls{ao} and shaping approaches diminishes as the \gls{ris} evolves from \gls{d} to fully-connected \gls{bd};
		\item The proposed geodesic \gls{rcg} method is efficient and the optimization cost of practically-sized \gls{bd}-\gls{ris} remains low;
		\item The solutions are robust to channel estimation errors and extendable to symmetric constraint with minimal degradation.
	\end{itemize}
	% \end{subsection}

	\emph{Notation:}
	Italic, bold lower-case, and bold upper-case letters indicate scalars, vectors and matrices, respectively.
	$\jmath$ denotes the imaginary unit.
	$\mathbb{R}$ and $\mathbb{C}$ denote the set of real and complex numbers, respectively.
	$\mathbb{H}^{n \times n}$, $\mathbb{H}_+^{n \times n}$, $\mathbb{U}^{n \times n}$, and $\mathbb{P}^{n \times n}$ denote the set of $n \times n$ Hermitian, positive semi-definite, unitary, and permutation matrices, respectively.
	$\mathbf{0}$ and $\mathbf{I}$ are the zero and identity matrices with appropriate size, respectively.
	$\Re\{\cdot\}$ takes the real part of a complex number.
	$\mathbb{E}\{\cdot\}$ is the expectation operator.
	$\conv\{\cdot\}$ returns the convex hull of arguments.
	$\tr(\cdot)$ and $\det(\cdot)$ evaluate the trace and determinant of a square matrix, respectively.
	$\diag(\cdot)$ constructs a square matrix with arguments on the main (block) diagonal and zeros elsewhere.
	$\sv(\cdot)$, $\ran(\cdot)$, and $\ker(\cdot)$ evaluate the singular values, range, and kernel of a matrix, respectively.
	% $\sv(\cdot)$ extracts the singular values of a matrix as a non-ascending vector.
	% $\card(\cdot)$ returns the cardinality of a set.
	$\mathrm{vec}(\cdot)$ stacks the columns of a matrix as a vector.
	$\lvert \cdot \rvert$, $\lVert \cdot \rVert$, and $\lVert \cdot \rVert _\mathrm{F}$ denote the absolute value, Euclidean norm, and Frobenius norm, respectively.
	$\sigma_n(\cdot)$ and $\lambda_n(\cdot)$ are the $n$-th largest singular value and eigenvalue, respectively.
	% $\boldsymbol{\sigma}(\cdot)$ and $\boldsymbol{\lambda}(\cdot)$ are the corresponding vectors.
	$(\cdot)^*$, $(\cdot)^\mathsf{T}$, $(\cdot)^\mathsf{H}$, $(\cdot)^\dagger$, $(\cdot)^{\star}$ denote the conjugate, transpose, conjugate transpose (Hermitian), Moore-Penrose inverse, and stationary point, respectively.
	$[N]$ is a shortcut for $\{1,2,\ldots,N\}$.
	$(\cdot)_{[x:y]}$ is a shortcut for $(\cdot)_x,(\cdot)_{x+1},\ldots,(\cdot)_y$.
	$\odot$ denotes the Hadamard product.
	$\mathcal{O}(\cdot)$ is the big-O notation.
	$\mathcal{N}_{\mathbb{C}}(\mathbf{0}, \mathbf{\Sigma})$ is the multivariate \gls{cscg} distribution with mean $\mathbf{0}$ and covariance $\mathbf{\Sigma}$.
	$\sim$ means ``distributed as''.
\end{section}

>>>>

	Although above works offered inspiring glimpses into the channel shaping potential of passive \gls{ris}, they neither provided in-depth theoretical analysis nor characterized the achievable singular value region.
	Most works \cite{ElMossallamy2021,Meng2023,Huang2023,Chae2023,Wu2019,Santamaria2023a} have also been confined to the conventional \gls{d} architecture where each \gls{ris} element is connected to a dedicated impedance and behaves independently of the others, namely, the wave impinging on one element is entirely scattered by itself.
	This architecture is modeled by a diagonal scattering matrix with unit-magnitude diagonal entries that, in the ideal case, applies a phase shift to the incident signal.
	The idea was soon extended to \gls{bd}-\gls{ris} with group-connected architecture that connects elements within the same group via passive reconfigurable circuit components \cite{Shen2020a} that can be symmetric (e.g., capacitors and inductors) or asymmetric (e.g, ring hybrids and branch-line hybrids \cite{Ahn2006}).
	As such, the wave impinging on one element is able to propagate within the circuit and depart partially from the others of the same group.
	It can thus manipulate both amplitude and phase of the scattered wave while remaining globally passive.
	The main manufacturing complexity of \gls{bd}-\gls{ris} lies in the design and implementation of the circuit network.
	Fortunately, topologies such as tree- and forest-connected graphs have been proposed to reduce the number of components for a flexible cost-performance tradeoff \cite{Nerini2024}.
	Other practical challenges such as channel estimation \cite{Li2024}, mutual coupling \cite{Li2023f}, wideband modelling \cite{Li2024a}, multi-sector coverage \cite{Li2023c}, and hardware implementation \cite{Tapie25e} have also been studied in recent literature.
	\gls{bd}-\gls{ris} has been proved to achieve higher spectral efficiency than \gls{d}-\gls{ris} and higher energy efficiency than active \gls{ris} and \gls{af} relay \cite{Santamaria2023,Fang2023,Zhou2023}.
	However, the interplay between \gls{bd}-\gls{ris} and \gls{mimo} is still at infancy stage and its channel shaping potential remains largely unexplored.
	% Most relevant works are confined to the scenario
	For example, the rate maximization problem \cite{Bartoli2023} has only been tacked in the special case where the direct channel is negligible and the \gls{bd}-\gls{ris} is fully-connected.
	Under those conditions, the mathematical modeling of \gls{bd}-\gls{ris} coincides with that of \gls{af} relay of unit power, although the operation mechanism and noise characteristics are distinct.

	When it comes to signal processing, existing works have mainly invoked the {quasi-Newton} method \cite{Shen2020a}, the \gls{pdd} method \cite{Zhou2023}, and the generic (i.e., non-geodesic) \gls{rcg} method \cite{Li2023c} for the optimization of \gls{bd}-\gls{ris}.
	The first solves an unconstrained problem and projects the solution back to the feasible domain without optimality guarantee.
	The second alternates between the primal variables in the inner layer and the penalty coefficient in the outer layer.
	It is often used to tackle coupled constraints (e.g., \gls{sinr} thresholds under active and passive beamforming) and can be computationally expensive (e.g., $\mathcal{O}(N_{\mathrm{s}}^2)$ for \gls{d}-\gls{ris} and $\mathcal{O}(N_{\mathrm{s}}^4)$ for fully-connected \gls{bd}-\gls{ris}) \cite[Table~I]{Zhou2023}.
	The third applies the conjugate gradient method on generic Riemannian manifolds.
	Each iteration consists of an addition on the tangent space and a retraction to the feasible domain, which constitutes a zigzag path departing from and returning to the manifold.
	However, none of them effectively exploits the special structure of \gls{bd}-\gls{ris} for accelerated convergence.

	% \begin{subsection}{Contributions}
	This paper is motivated by a fundamental question:
	\emph{What is the channel shaping capability, in terms of singular values and their functions, of a passive \gls{ris} in \gls{mimo} channels?}
	Unlike existing works that focus on specific performance metrics or deployment scenarios, we aim for an understanding of the theoretical shaping limits (via analysis) and the achievable shaping results (via optimization) that are broadly applicable across diverse wireless applications.
	We believe a comprehensive shaping answer can serve as a theoretical support/reference for the vast number of \gls{ris} research papers and real-world applications.
	Without a framework that identifies the fundamental shaping limits of \gls{ris}, the design of truly optimal and efficient architectures will remain elusive.
	The contributions of this paper are summarized below.

@@

	Fifth, we validate the analytical bounds and the numerical methods by extensive simulation.
	It is concluded that:
	\begin{itemize}
		\item \gls{bd}-\gls{ris} widens the dynamic range of channel singular values for enhanced rate, power, and physical-layer security;
		\item The shaping benefits of \gls{bd}-\gls{ris} over \gls{d}-\gls{ris} scale with the number of elements, group size, and \gls{mimo} dimensions;
		% \item The optimal shaping solutions for communication and power transfer coincide when direct channel is negligible and the \gls{bd}-\gls{ris} is fully-connected;
		% \item The optimally shaped \gls{mimo} channel for communication and power transfer\footnote{with \gls{rf} combining} coincide with each other
		% power-gain optimal, rate-optimal, and power-transfer \gls{bd}-\gls{ris} scattering matrix coincide
		% \item A fully-connected \gls{bd}-\gls{ris} may be designed in closed-form for simultaneous optimality of \gls{mimo} communication and power transfer;
		% \item \gls{bd}-\gls{ris} can activate multi-stream transmission and achieve the asymptotic \gls{dof} at lower transmit power than \gls{d}-\gls{ris};
		\item \gls{bd}-\gls{ris} can activate multi-stream transmission at much lower transmit power than \gls{d}-\gls{ris};
		\item The performance gap between the optimal joint beamforming and the low-complexity shaping solution diminishes as the \gls{ris} evolves from \gls{d} to fully-connected \gls{bd};
		\item The proposed geodesic \gls{rcg} method is efficient and the optimization cost of practically-sized \gls{bd}-\gls{ris} is affordable;
		\item The solutions are robust to channel estimation errors and extendable to symmetric case with minimal degradation.
	\end{itemize}
	% \end{subsection}

	\emph{Notation:}
	Italic, bold lower-case, and bold upper-case letters indicate scalars, vectors and matrices, respectively.
	$\jmath$ denotes the imaginary unit.
	$\mathbb{R}$ and $\mathbb{C}$ denote the set of real and complex numbers, respectively.
	$\mathbb{H}^{n \times n}$, $\mathbb{H}_+^{n \times n}$, $\mathbb{U}^{n \times n}$, and $\mathbb{P}^{n \times n}$ denote the set of $n \times n$ Hermitian, positive semi-definite, unitary, and permutation matrices, respectively.
	$\mathbf{0}$ and $\mathbf{I}$ are the zero and identity matrices with appropriate size, respectively.
	$\Re\{\cdot\}$ takes the real part of a complex number.
	$\mathbb{E}\{\cdot\}$ is the expectation operator.
	$\conv\{\cdot\}$ returns the convex hull of arguments.
	$\tr(\cdot)$ and $\det(\cdot)$ evaluate the trace and determinant of a square matrix, respectively.
	$\diagm(\cdot)$ constructs a square matrix with arguments on the main (block) diagonal and zeros elsewhere.
	$\sv(\cdot)$, $\ran(\cdot)$, and $\ker(\cdot)$ evaluate the singular values, range, and kernel of a matrix, respectively.
	% $\sv(\cdot)$ extracts the singular values of a matrix as a non-ascending vector.
	% $\card(\cdot)$ returns the cardinality of a set.
	$\mathrm{vec}(\cdot)$ stacks the columns of a matrix as a vector.
	$\lvert \cdot \rvert$, $\lVert \cdot \rVert$, and $\lVert \cdot \rVert _\mathrm{F}$ denote the absolute value, Euclidean norm, and Frobenius norm, respectively.
	$\sigma_n(\cdot)$ and $\lambda_n(\cdot)$ are the $n$-th largest singular value and eigenvalue, respectively.
	% $\boldsymbol{\sigma}(\cdot)$ and $\boldsymbol{\lambda}(\cdot)$ are the corresponding vectors.
	$(\cdot)^*$, $(\cdot)^\mathsf{T}$, $(\cdot)^\mathsf{H}$, $(\cdot)^\dagger$, $(\cdot)^{\star}$ denote the conjugate, transpose, conjugate transpose (Hermitian), Moore-Penrose inverse, and stationary point, respectively.
	$[N]$ is a shortcut for $\{1,2,\ldots,N\}$.
	$(\cdot)_{[x:y]}$ is a shortcut for $(\cdot)_x,(\cdot)_{x+1},\ldots,(\cdot)_y$.
	$\odot$ denotes the Hadamard product.
	$\mathcal{O}(\cdot)$ is the big-O notation.
	$\mathcal{N}_{\mathbb{C}}(\mathbf{0}, \mathbf{\Sigma})$ is the multivariate \gls{cscg} distribution with mean $\mathbf{0}$ and covariance $\mathbf{\Sigma}$.
	$\sim$ means ``distributed as''.
\end{section}
